---
title: "a2_task2_dani_sclafani"
author: "Danielle Sclafani"
date: "2/6/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE, warnings = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
library(kableExtra)

```


```{r, warning=FALSE}
palmetto <- read_csv(here("data", "palmetto.csv")) %>% 
  clean_names() %>% 
  select(height, length, width, green_lvs, species) %>% 
  mutate(species_name = case_when(
    species == 1 ~ "Serenoa repens",
    species == 2 ~ "Sabal etonia"
  ))
```

## Data Visualizations

```{r, warning = FALSE, fig.cap= "Figure 1.0: Distribution of _S. etonia_ and _S. repens_ canopy heights (cm)"}
# data visualization of height comparison

ggplot(data = palmetto, aes(x = height, y = width)) + geom_point(aes(color = species_name), alpha = 0.7) +
                                                          scale_color_manual(values = c("cyan4", "blue"))  +
  theme_minimal()+
  labs(x = "Height (cm)", y ="Width (cm)", color = "Species")
```
The two species of trees, _S. etonia_ and _S. repens_ have very similar distributions of height (cm) and width (cm), however, there seems to be more _S. repens_ trending towards a smaller width and _S. etonia_ having a larger width.

```{r, warning = FALSE, fig.cap= "Figure 2.0: Comparison of canopy Length (cm) of _S. etonia_ and _S. repens_"}
ggplot(data = palmetto, aes(x = length)) + geom_boxplot(aes(fill = species_name), alpha = 0.7) +
                                                          scale_fill_manual(values = c("pink", "orange")) +
  theme_minimal() +
  labs(x = "Canopy Length (cm)", fill = "Species")
```

The overall distribution of _S. entonia_ and _S. repens_ overlap pretty significantly however, the mean canopy length (cm) of _S. entonia_ is slightly larger than _S. repens_. 


```{r, warning = FALSE, fig.cap="Figure 3: Comparison of the number of green leaves present on _S. etonia_ and _S. repens_ trees."}

# data visualization exploring number of green leaves on each species

ggplot(data = palmetto, aes(x = green_lvs)) + geom_histogram(aes(fill = species_name), alpha = 0.7) + 
  scale_fill_manual(values = c("green", "darkgreen")) +
  theme_minimal()+
  labs(x = "Number of Green Leaves", fill = "Species", y = " ")
```

The distribution of the number of green leaves on each tree species appears to provide some insight into how the trees may be identified. _S. etonia_ appears to have a lower mean number of leaves than _S. repens_. The distribution of the number of leaves for _S. repens_ is larger than that of _S. etonia_. 


## Binary Logistic Regression


```{r, warning = FALSE}
#made the species into factors, Sabal etonia = 0, Serenoa repens = 1, because Serenoa repens is the 1 factor, the regression model will show us the probability that it is Serenoa repens

palm_fact <- palmetto %>% 
  mutate(species_name = fct_drop(species_name))

#making the binary logistic regression
palm_blr <- glm(species_name ~ height + length + green_lvs + width,
                data = palm_fact,
                family = "binomial")

# looking at the outcomes

summary(palm_blr)

```

Table 1.0: Results of the Binary Logistics Regression model. The estimate column contains the coefficient for each variable, which states as each variable increases by one unit, the probability of the tree being _S. repens_ increases or decreases by the amount of the coefficient. 
```{r, warning = FALSE}
# making a table of blr results

tidy_palm_blr <- broom::tidy(palm_blr)
tidy_palm_blr

kable(tidy_palm_blr, caption = "Palmetto Binary Logistic Regression Results") %>% 
  kable_classic_2(full_width = F) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

```{r}
# getting the probabilities, probability that something is a S. repens is in the .fitted column
blr_fitted <- palm_blr %>% 
  broom::augment(type.predict = "response")
```

```{r}
# adding a column that says the species the observation should be based on the model
palm_prob <- blr_fitted %>% 
  mutate(predicted_species = case_when(
    .fitted >= 0.5 ~ "Serenoa repens",
    .fitted < 0.5 ~ "Sabal etonia"
  ))

#using case when
palm_accurate <- palm_prob %>% 
  mutate(accurate = case_when(
    species_name == predicted_species ~ "Accurate",
    species_name != predicted_species ~ "Inaccurate")) 

#palm_accurate table
palm_table <- palm_accurate %>% 
  group_by(species_name, accurate) %>% 
  count()


```

Table 2.0 Displays the accuracy of the Binary Logistic Model. The percent of each species accurately and inaccurately predicted by the model are provided. The counts of accurate and inaccurate predictions are also included in parantheses. 
```{r, warning = FALSE}
#percent attempt
percent_attempt <- palm_table %>%
  pivot_wider(species_name, names_from = "accurate",
              values_from = "n") %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns() %>% 
  rename(Species = species_name)

percent_attempt %>% 
  kbl(caption = "Percentage and Occurence of Binary Logistic Model predicting the species Accurately and Inaccurately") %>% 
  kable_classic_2(full_width = F, html_font = "Cambria")

```


Data Citation: Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5
