---
title: "a2_task2_dani_sclafani"
author: "Danielle Sclafani"
date: "2/6/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE, warnings = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
library(kableExtra)

```


```{r, warning=FALSE}
palmetto <- read_csv(here("data", "palmetto.csv")) %>% 
  clean_names() %>% 
  select(height, length, width, green_lvs, species) %>% 
  mutate(species_name = case_when(
    species == 1 ~ "Serenoa repens",
    species == 2 ~ "Sabal etonia"
  ))
```

## Data Visualizations

```{r, warning = FALSE, fig.cap= "Figure 1.0: Distribution of _S. etonia_ and _S. repens_ canopy heights (cm)"}
# data visualization of height comparison

ggplot(data = palmetto, aes(x = height, y = width)) + geom_point(aes(color = species_name), alpha = 0.7) +
                                                          scale_color_manual(values = c("cyan4", "blue"))  +
  theme_minimal()+
  labs(x = "Height (cm)", y ="Width (cm)", color = "Species")
```
The two species of trees, _S. etonia_ and _S. repens_ have very similar distributions of height (cm) and width (cm), however, there seems to be more _S. repens_ trending towards a smaller width and _S. etonia_ having a larger width.

```{r, warning = FALSE, fig.cap= "Figure 2.0: Comparison of canopy Length (cm) of _S. etonia_ and _S. repens_"}
ggplot(data = palmetto, aes(x = length)) + geom_boxplot(aes(fill = species_name), alpha = 0.7) +
                                                          scale_fill_manual(values = c("pink", "orange")) +
  theme_minimal() +
  labs(x = "Canopy Length (cm)", fill = "Species")
```

The overall distribution of _S. entonia_ and _S. repens_ overlap pretty significantly however, the mean canopy length (cm) of _S. entonia_ is slightly larger than _S. repens_. 


```{r, warning = FALSE, fig.cap="Figure 3: Comparison of the number of green leaves present on _S. etonia_ and _S. repens_ trees."}

# data visualization exploring number of green leaves on each species

ggplot(data = palmetto, aes(x = green_lvs)) + geom_histogram(aes(fill = species_name), alpha = 0.7) + 
  scale_fill_manual(values = c("green", "darkgreen")) +
  theme_minimal()+
  labs(x = "Number of Green Leaves", fill = "Species", y = " ")
```

The distribution of the number of green leaves on each tree species appears to provide some insight into how the trees may be identified. _S. etonia_ appears to have a lower mean number of leaves than _S. repens_. The distribution of the number of leaves for _S. repens_ is larger than that of _S. etonia_. 


## Binary Logistic Regression


```{r, warning = FALSE}
#made the species into factors, Sabal etonia = 0, Serenoa repens = 1, because Serenoa repens is the 1 factor, the regression model will show us the probability that it is Serenoa repens

palm_fact <- palmetto %>% 
  mutate(species_name = fct_drop(species_name))

#making the binary logistic regression
palm_blr <- glm(species_name ~ height + length + green_lvs + width,
                data = palm_fact,
                family = "binomial")

# looking at the outcomes

summary(palm_blr)

```

```{r}
# making a table of blr results

tidy_palm_blr <- broom::tidy(palm_blr)
tidy_palm_blr

kable(tidy_palm_blr, caption = "Palmetto Binary Logistic Regression Results") %>% 
  kable_classic_2(full_width = F) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

```{r}
# getting the probabilities, probability that something is a S. repens is in the .fitted column
blr_fitted <- palm_blr %>% 
  broom::augment(type.predict = "response")
```

```{r}
# adding a column that says the species the observation should be based on the model
palm_prob <- blr_fitted %>% 
  mutate(predicted_species = case_when(
    .fitted >= 0.5 ~ "Serenoa repens",
    .fitted < 0.5 ~ "Sabal etonia"
  ))

#using case when?
palm_accurate <- palm_prob %>% 
  mutate(accurate = case_when(
    species_name == predicted_species ~ "Accurate",
    species_name != predicted_species ~ "Inaccurate"))

#uniting the two columns
unite_palm <- palm_prob %>% 
  select(species_name, predicted_species) %>% 
  unite(accurate, species_name : predicted_species, sep = ",", remove = FALSE)

#counting the amount that are accurate and false
count_attempt <- unite_palm %>% 
  group_by(accurate) %>% 
  count()

#renaming for the table
table <- count_attempt %>% 
  mutate(Results = case_when(
    accurate == "Sabal etonia,Sabal etonia" ~ "Correctly identified as Sabal etonia",
    accurate == "Sabal etonia,Serenoa repens" ~ "Incorrectly identified as Serenoa repens",
    accurate == "Serenoa repens,Serenoa repens" ~ "Correctly identified as Serenoa repens",
    accurate == "Serenoa repens,Sabal etonia" ~ "Incorrectly identified as Sabal etonia"
  ))

#renaming table
table_2 <- table %>% 
  select(Results, n)
```


Data Citation: Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5
